---
title: "Assigment7 - Working with XML and JSON in R"
author: "Abdelmalek Hajjam"
date: "10/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Pick three of your favorite books on one of your favorite subjects. At least one of the books should have more
than one author. For each book, include the title, authors, and two or three other attributes that you find
interesting.
Take the information that you’ve selected about these three books, and separately create three files which
store the book’s information in HTML (using an html table), XML, and JSON formats (e.g. “books.html”,
“books.xml”, and “books.json”). To help you better understand the different file structures, I’d prefer that you
create each of these files “by hand” unless you’re already very comfortable with the file formats.
Write R code, using your packages of choice, to load the information from each of the three sources into
separate R data frames. Are the three data frames identical?
Your deliverable is the three source files and the R code. If you can, package your assignment solution up into
an .Rmd file and publish to rpubs.com. [This will also require finding a way to make your three text files
accessible from the web].**


```{r results='hide', message=FALSE, warning=FALSE}
library(stringr)
library(XML) #for xml files processing
library(RCurl)
library(rvest)  # my favorite
library(dplyr)
library(jsonlite) # for json files processing
library("kableExtra") # for beautify dataframes display


# to get rid of the scientific notation for ISBN long number and leave it as character, we include the folling option:
options(scipen=999) 
#without this option on, the ISBN of our books will be presented in a scientific notation, i.e. with exponential display
```


## Loading from an HTML file that has an html table inside it
```{r message=FALSE, warning=FALSE}
url <- "https://raw.githubusercontent.com/theoracley/Data607/master/mybooks.html"

#first method
# theURL <- getURL(url)
# myTable <- readHTMLTable(theURL, header=TRUE, which=1)
# myTable


# #second Method (more work)
# page <- read_html(url)
# tables <- html_nodes(page, "table")
# myTable <- html_table(tables[1], fill = TRUE)
# myTable[[1]]

#third method
myTable <- read_html(url)
htmlBookDS <- myTable %>%
  html_nodes("table") %>%
  .[[1]] %>%
  html_table(fill = NA) # html_table converts data to an R data frame automatically

#htmlBookDS
#beautify the display
htmlBookDS  %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  scroll_box(width = "100%", height = "200px")
 
```


## Loading from an xml file
```{r message=FALSE, warning=FALSE}

url <- "https://raw.githubusercontent.com/theoracley/Data607/master/mybooks.xml"
 
xmlData <- read_xml(url)
aiBooks <- xml_children(xmlData)

xmlBookDS <- c()
for (i in 1:length(aiBooks)){
  xmlBookDS <- rbind(xmlBookDS,xml_text(xml_children(aiBooks[i])))
}

xmlBookDS <- data.frame(xmlBookDS)

colnames(xmlBookDS) <- xml_name(xml_children(aiBooks[1]))
#xmlBookDS

#beautify the display
xmlBookDS  %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  scroll_box(width = "100%", height = "200px")

```

## Loading from a json file
```{r message=FALSE, warning=FALSE}

url <- "https://raw.githubusercontent.com/theoracley/Data607/master/mybooks.json"

#fromJSON method converts any json object to an R data frame.
jsonBookDS <- fromJSON(url)
#jsonBookDS

#beautify the display
jsonBookDS  %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  scroll_box(width = "100%", height = "200px")


```


## Conclusion:
**As we can see, all datasets created from html, xml and json files have exactly the same resulting dataframes. While all 3 files have different structures , they all yield the same resulting dataframes when choosing special package for each file type.**
